---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

## workboots

**Author:** [Mark Rieke](https://www.thedatadiary.net/about/) <br/>
**License:** [MIT](https://github.com/markjrieke/workboots/blob/main/LICENSE)

<!-- badges: start -->
[![R-CMD-check](https://github.com/markjrieke/workboots/workflows/R-CMD-check/badge.svg)](https://github.com/markjrieke/workboots/actions)
<!-- badges: end -->

### Overview

`{workboots}` is a tidy method of generating bootstrap prediction intervals for arbitrary model types from a tidymodel workflow. 

By using [bootstrap resampling](https://en.wikipedia.org/wiki/Bootstrapping_(statistics)), we can create many models --- one for each resample. Each model will be slightly different based on the resample it was trained on. Each model will also generate slightly different predictions for new data, allowing us to generate a prediction distribution for models that otherwise just return point predictions.

### Installation

Currently, this package lives exclusively on github, though work is in progress to publish to CRAN. You can install the development version with the [devtools](https://cran.r-project.org/package=devtools) or [remotes](https://cran.r-project.org/package=remotes) package:

```{r, eval = FALSE}
devtools::install_github("markjrieke/workboots")
```

### Usage

`{workboots}` builds on top of the [`{tidymodels}`](https://www.tidymodels.org/) suite of packages. Teaching how to use `{tidymodels}` is beyond the scope of this package, but some helpful resources are linked at the bottom of this README. 

To get started, we'll need to create a workflow.

```{r}
library(tidymodels)

# load our dataset
data("penguins")
penguins <- penguins %>% drop_na()

# split data into testing & training sets
set.seed(123)
penguins_split <- initial_split(penguins)
penguins_test <- testing(penguins_split)
penguins_train <- training(penguins_split)

# create a workflow
penguins_wf <- 
  workflow() %>%
  add_recipe(recipe(body_mass_g ~ ., data = penguins_train) %>% step_dummy(all_nominal())) %>%
  add_model(boost_tree("regression") %>% set_engine("xgboost"))
```

XGBoost can only generate point predictions, but with `{workboots}` we can train many models to generate many predictions and create a summary with lower and upper bounds:

```{r}
library(workboots)

penguins_preds <-
  penguins_wf %>%
  predict_boots(
    n = 100,
    training_data = penguins_train,
    new_data = penguins_test
  )

penguins_preds %>%
  summarise_predictions()
```

### Bug reports/feature requests 

This package is still in its infancy --- if you notice a bug, want to request a new feature, or have recommendations on improving documentation, please [open an issue](https://github.com/markjrieke/workboots/issues) in this repository.

### Tidymodels Resources

* [Getting started with Tidymodels](https://www.tidymodels.org/start/)
* [Tidy Modeling with R](https://www.tmwr.org/)
* [Julia Silge's Blog](https://juliasilge.com/blog/) provides use cases of tidymodels with weekly [#tidytuesday](https://github.com/rfordatascience/tidytuesday) datasets. 
